---
title: "Transformations"
output: html_document
---

**CARBOHYDRATE:** DR is log; C is log; HS is untransformed

**LIPID:**  DR, C and HS untransformed;

**GLYCOGEN:** DR, C and HS all square root

**PROTEIN:** DR,C and HS is  untransformed

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(cowplot)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(plyr)
library(forcats)
setwd("C:/Users/Anna/Dropbox/AMP_EGK/Projects/HalfSib/scripts")

combined<- read.table("../processed_data/perfly_overall_dataset.txt", sep="\t",header=TRUE,stringsAsFactors=FALSE)

all_prot<-read.table("../processed_data/all_prot_tot.txt", sep="\t",header=TRUE,stringsAsFactors=FALSE)
all_carb<-read.table("../processed_data/all_carb_tot.txt", sep="\t",header=TRUE,stringsAsFactors=FALSE)
all_gly<-read.table("../processed_data/all_gly_tot.txt", sep="\t",header=TRUE,stringsAsFactors=FALSE)
all_lip<-read.table("../processed_data/all_lip_tot.txt", sep="\t",header=TRUE,stringsAsFactors=FALSE)

which(all_lip$hs_lip <0) #1
all_lip$hs_lip[all_lip$hs_lip<0] <- NA
write.table(all_lip, "../processed_data/all_lip_tot.txt", sep="\t")
```

```{r}
test<-all_carb %>% separate(RIL, c("damid", "rep"))
testy<-cbind(all_carb$RIL, test)
testy$RIL<-all_carb$RIL
test_A<-testy[,-1]
sireid<-rep(1:72, each=3)
damid<-c(1:216)
sire_dam_sets<- cbind(sireid,damid)
dat_carb_sire <- merge(test_A,sire_dam_sets,by="damid")
dat_carb_sire$sireid <-as.factor(dat_carb_sire$sireid)
dat_carb_sire$damid <-as.factor(dat_carb_sire$damid)

c_carb<- dplyr::summarize(group_by(dat_carb_sire,sireid),c_carb_sm=mean(c_carb, na.rm = TRUE))#
dr_carb<- dplyr::summarize(group_by(dat_carb_sire,sireid),dr_carb_sm=mean(dr_carb, na.rm = TRUE))#
hs_carb<- dplyr::summarize(group_by(dat_carb_sire,sireid),hs_carb_sm=mean(hs_carb, na.rm = TRUE))#

carb_<-left_join(c_carb, dr_carb, by="sireid")
all_carb<-left_join(carb_, hs_carb, by="sireid")

names(all_carb)[names(all_carb) == "c_carb_sm"] <- "c_carb"
names(all_carb)[names(all_carb) == "dr_carb_sm"] <- "dr_carb"
names(all_carb)[names(all_carb) == "hs_carb_sm"] <- "hs_carb"

write.table(all_carb, "../processed_data/carb_sire_avg.txt", sep="\t")
```

```{r}
test<-all_lip %>% separate(RIL, c("damid", "rep"))
testy<-cbind(all_lip$RIL, test)
testy$RIL<-all_lip$RIL
test_A<-testy[,-1]
sireid<-rep(1:72, each=3)
damid<-c(1:216)
sire_dam_sets<- cbind(sireid,damid)
dat_lip_sire <- merge(test_A,sire_dam_sets,by="damid")
dat_lip_sire$sireid <-as.factor(dat_lip_sire$sireid)
dat_lip_sire$damid <-as.factor(dat_lip_sire$damid)

c_lip<- dplyr::summarize(group_by(dat_lip_sire,sireid),c_lip_sm=mean(c_lip, na.rm = TRUE))#
dr_lip<- dplyr::summarize(group_by(dat_lip_sire,sireid),dr_lip_sm=mean(dr_lip, na.rm = TRUE))#
hs_lip<- dplyr::summarize(group_by(dat_lip_sire,sireid),hs_lip_sm=mean(hs_lip, na.rm = TRUE))#

lip_<-left_join(c_lip, dr_lip, by="sireid")
all_lip<-left_join(lip_, hs_lip, by="sireid")

names(all_lip)[names(all_lip) == "c_lip_sm"] <- "c_lip"
names(all_lip)[names(all_lip) == "dr_lip_sm"] <- "dr_lip"
names(all_lip)[names(all_lip) == "hs_lip_sm"] <- "hs_lip"

write.table(all_lip, "../processed_data/lip_sire_avg.txt", sep="\t")
```

```{r}
test<-all_prot %>% separate(RIL, c("damid", "rep"))
testy<-cbind(all_prot$RIL, test)
testy$RIL<-all_prot$RIL
test_A<-testy[,-1]
sireid<-rep(1:72, each=3)
damid<-c(1:216)
sire_dam_sets<- cbind(sireid,damid)
dat_prot_sire <- merge(test_A,sire_dam_sets,by="damid")
dat_prot_sire$sireid <-as.factor(dat_prot_sire$sireid)
dat_prot_sire$damid <-as.factor(dat_prot_sire$damid)

c_prot<- dplyr::summarize(group_by(dat_prot_sire,sireid),c_prot_sm=mean(c_prot, na.rm = TRUE))#
dr_prot<- dplyr::summarize(group_by(dat_prot_sire,sireid),dr_prot_sm=mean(dr_prot, na.rm = TRUE))#
hs_prot<- dplyr::summarize(group_by(dat_prot_sire,sireid),hs_prot_sm=mean(hs_prot, na.rm = TRUE))#

prot_<-left_join(c_prot, dr_prot, by="sireid")
all_prot<-left_join(prot_, hs_prot, by="sireid")

names(all_prot)[names(all_prot) == "c_prot_sm"] <- "c_prot"
names(all_prot)[names(all_prot) == "dr_prot_sm"] <- "dr_prot"
names(all_prot)[names(all_prot) == "hs_prot_sm"] <- "hs_prot"

write.table(all_prot, "../processed_data/prot_sire_avg.txt", sep="\t")
```

```{r}
test<-all_gly %>% separate(RIL, c("damid", "rep"))
testy<-cbind(all_gly$RIL, test)
testy$RIL<-all_gly$RIL
test_A<-testy[,-1]
sireid<-rep(1:72, each=3)
damid<-c(1:216)
sire_dam_sets<- cbind(sireid,damid)
dat_gly_sire <- merge(test_A,sire_dam_sets,by="damid")
dat_gly_sire$sireid <-as.factor(dat_gly_sire$sireid)
dat_gly_sire$damid <-as.factor(dat_gly_sire$damid)

c_gly<- dplyr::summarize(group_by(dat_gly_sire,sireid),c_gly_sm=mean(c_gly, na.rm = TRUE))#
dr_gly<- dplyr::summarize(group_by(dat_gly_sire,sireid),dr_gly_sm=mean(dr_gly, na.rm = TRUE))#
hs_gly<- dplyr::summarize(group_by(dat_gly_sire,sireid),hs_gly_sm=mean(hs_gly, na.rm = TRUE))#

gly_<-left_join(c_gly, dr_gly, by="sireid")
all_gly<-left_join(gly_, hs_gly, by="sireid")

names(all_gly)[names(all_gly) == "c_gly_sm"] <- "c_gly"
names(all_gly)[names(all_gly) == "dr_gly_sm"] <- "dr_gly"
names(all_gly)[names(all_gly) == "hs_gly_sm"] <- "hs_gly"

write.table(all_gly, "../processed_data/gly_sire_avg.txt", sep="\t")
```


```{r}
#prot untransformed
all_prot$c_prot_T<-(all_prot$c_prot) 
all_prot$dr_prot_T<-(all_prot$dr_prot)
all_prot$hs_prot_T<-(all_prot$hs_prot) 

testing<-select(all_prot, sireid, c_prot_T, dr_prot_T, hs_prot_T)
write.table(testing, "../processed_data/prot_transform_sire_avg.txt", sep="\t")


#lip untransformed
all_lip$c_lip_T<-(all_lip$c_lip) 
all_lip$dr_lip_T<-(all_lip$dr_lip) 
all_lip$hs_lip_T<-(all_lip$hs_lip) 

testing<-select(all_lip, sireid, c_lip_T, dr_lip_T, hs_lip_T)
write.table(testing, "../processed_data/lipid_transform_sire_avg.txt", sep="\t")

#carb
all_carb$c_carb_T<-log(all_carb$c_carb) 
all_carb$dr_carb_T<-log(all_carb$dr_carb) 
all_carb$hs_carb_T<-log(all_carb$hs_carb) 
testing<-select(all_carb, sireid, c_carb_T, dr_carb_T, hs_carb_T)
write.table(testing, "../processed_data/carbs_transform_sire_avg.txt", sep="\t")

#glycogne: dr and c untransfomred, hs square root
all_gly$c_gly_T<-(all_gly$c_gly) 
all_gly$dr_gly_T<-(all_gly$dr_gly) 
all_gly$hs_gly_T<-(all_gly$hs_gly^1/2) 

test<-select(all_gly, sireid, c_gly_T, dr_gly_T, hs_gly_T)
write.table(test, "../processed_data/glycogen_transform_sire_avg.txt", sep="\t")

```


#Carbohydrate
**-DR: sq root
-C: log
-HS: untransformed**

_Yes, it looks like there are outliers, these are samples that were paired w their technical replicate so not actually an "outlier"_
```{r}
#high sugar (log)
p1a<-ggplot(data=all_carb, aes(x = (hs_carb))) +
  geom_histogram(bins=20) 
p1b<-ggplot(data=all_carb, aes(x = (hs_carb)^(1/2))) +
  geom_histogram(bins=20) 
p1c<-ggplot(data=all_carb, aes(x = log(hs_carb))) +
  geom_histogram(bins=20) 
p1d<-ggplot(data=all_carb, aes(x = (hs_carb)^(2))) +
  geom_histogram(bins=20) 
plot_grid(p1a,p1b,p1c,p1d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#control (log)
p2a<-ggplot(data=all_carb, aes(x = (c_carb))) +
  geom_histogram(bins=20)
p2b<-ggplot(data=all_carb, aes(x = (c_carb)^(1/2))) +
  geom_histogram(bins=20) 
p2c<-ggplot(data=all_carb, aes(x = log(c_carb))) +
  geom_histogram(bins=20)
p2d<-ggplot(data=all_carb, aes(x = (c_carb)^(2))) +
  geom_histogram(bins=20) 
plot_grid(p2a,p2b,p2c,p2d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#dr (log)
p3a<-ggplot(data=all_carb, aes(x = (dr_carb))) +
  geom_histogram(bins=20) 
p3b<-ggplot(data=all_carb, aes(x = (dr_carb)^(1/2))) +
  geom_histogram(bins=20) 
p3c<-ggplot(data=all_carb, aes(x = log(dr_carb))) +
  geom_histogram(bins=20) 
p3d<-ggplot(data=all_carb, aes(x = (dr_carb)^(2))) +
  geom_histogram(bins=20) 
plot_grid(p3a,p3b,p3c,p3d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#DR sq root and C log, HS is untransformed
all_carb$c_carb_T<-log(all_carb$c_carb) 
all_carb$dr_carb_T<-log(all_carb$dr_carb) 
all_carb$hs_carb_T<-(all_carb$hs_carb) 
testing<-select(all_carb, RIL, c_carb_T, dr_carb_T, hs_carb_T)
write.table(testing, "../processed_data/carbs_transform.txt", sep="\t")

```

##GLYCOGEN
**-hs: square root
-c: square root
-dr: square root**
```{r glycogen}
#HS (square root)
p4a<-ggplot(data=all_gly, aes(x = (hs_gly))) +
  geom_histogram(bins=30) 
p4b<-ggplot(data=all_gly, aes(x = (hs_gly)^(1/2))) +
  geom_histogram(bins=30) 
p4c<-ggplot(data=all_gly, aes(x = log(hs_gly))) +
  geom_histogram(bins=30) 
p4d<-ggplot(data=all_gly, aes(x = (hs_gly)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p4a,p4b,p4c,p4d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#control (untransfomred)
p5a<-ggplot(data=all_gly, aes(x = (c_gly))) +
  geom_histogram(bins=30) 
p5b<-ggplot(data=all_gly, aes(x = (c_gly)^(1/2))) +
  geom_histogram(bins=30) 
p5c<-ggplot(data=all_gly, aes(x =log(c_gly))) +
  geom_histogram(bins=30) 
p5d<-ggplot(data=all_gly, aes(x = (c_gly)^(2))) +
  geom_histogram(bins=30)
plot_grid(p5a,p5b,p5c,p5d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#DR is untransformed
p6a<-ggplot(data=all_gly, aes(x = (dr_gly))) +
  geom_histogram(bins=30) 
p6b<-ggplot(data=all_gly, aes(x = (dr_gly)^(1/2))) +
  geom_histogram(bins=30) 
p6c<-ggplot(data=all_gly, aes(x =log(dr_gly))) +
  geom_histogram(bins=30) 
p6d<-ggplot(data=all_gly, aes(x = (dr_gly)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p6a,p6b,p6c,p6d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))


all_gly$c_gly_T<-(all_gly$c_gly^1/2) 
all_gly$dr_gly_T<-(all_gly$dr_gly^1/2) 
all_gly$hs_gly_T<-(all_gly$hs_gly^1/2) 

test<-select(all_gly, RIL, c_gly_T, dr_gly_T, hs_gly_T)
write.table(test, "../processed_data/glycogen_transform.txt", sep="\t")
```


##PROTEIN
**-hs: untransformed
-c: (log most likely, possibly sq root)
-dr: square root**
```{r protien}
#HS (untransfomred)
p7a<-ggplot(data=all_prot, aes(x = (hs_prot))) +
  geom_histogram(bins=30) 
p7b<-ggplot(data=all_prot, aes(x = (hs_prot)^(1/2))) +
  geom_histogram(bins=30) 
p7c<-ggplot(data=all_prot, aes(x = log(hs_prot))) +
  geom_histogram(bins=30) 
p7d<-ggplot(data=all_prot, aes(x = (hs_prot)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p7a,p7b,p7c,p7d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#control (log most likely, possibly sq root)
p8a<-ggplot(data=all_prot, aes(x = (c_prot))) +
  geom_histogram(bins=30) 
p8b<-ggplot(data=all_prot, aes(x = (c_prot)^(1/2))) +
  geom_histogram(bins=30) 
p8c<-ggplot(data=all_prot, aes(x = log(c_prot))) +
  geom_histogram(bins=30) 
p8d<-ggplot(data=all_prot, aes(x = (c_prot)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p8a,p8b,p8c,p8d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))
#DR (square root)
p9a<-ggplot(data=all_prot, aes(x = (dr_prot))) +
  geom_histogram(bins=30) 
p9b<-ggplot(data=all_prot, aes(x = (dr_prot)^(1/2))) +
  geom_histogram(bins=30) 
p9c<-ggplot(data=all_prot, aes(x = log(dr_prot))) +
  geom_histogram(bins=30) 
p9d<-ggplot(data=all_prot, aes(x = (dr_prot)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p9a,p9b,p9c,p9d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#hs: untransformed, Control,  (log most likely, possibly sq root); dr: square root

all_prot$c_prot_T<-(all_prot$c_prot) 
all_prot$dr_prot_T<-(all_prot$dr_prot)
all_prot$hs_prot_T<-(all_prot$hs_prot) 

testing<-select(all_prot, RIL, c_prot_T, dr_prot_T, hs_prot_T)
write.table(testing, "../processed_data/prot_transform.txt", sep="\t")

```

##LIPID
**-hs: untransformed 
-c: square root
-dr: untransformed**
```{r lipid}
#HS (untransformed)
p10a<-ggplot(data=all_lip, aes(x = (hs_lip))) +
  geom_histogram(bins=30) 
p10b<-ggplot(data=all_lip, aes(x = (hs_lip)^(1/2))) +
  geom_histogram(bins=30) 
p10c<-ggplot(data=all_lip, aes(x = log(hs_lip))) +
  geom_histogram(bins=30) 
p10d<-ggplot(data=all_lip, aes(x = (hs_lip)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p10a,p10b,p10c,p10d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#control (square root)
p11a<-ggplot(data=all_lip, aes(x = (c_lip))) +
  geom_histogram(bins=30) 
p11b<-ggplot(data=all_lip, aes(x = (c_lip)^(1/2))) +
  geom_histogram(bins=30) 
p11c<-ggplot(data=all_lip, aes(x = log(c_lip))) +
  geom_histogram(bins=30) 
p11d<-ggplot(data=all_lip, aes(x = (c_lip)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p11a,p11b,p11c,p11d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

#DR (untransformed)
p12a<-ggplot(data=all_lip, aes(x = (dr_lip))) +
  geom_histogram(bins=30) 
p12b<-ggplot(data=all_lip, aes(x = (dr_lip)^(1/2))) +
  geom_histogram(bins=30) 
p12c<-ggplot(data=all_lip, aes(x = log(dr_lip))) +
  geom_histogram(bins=30) 
p12d<-ggplot(data=all_lip, aes(x = (dr_lip)^(2))) +
  geom_histogram(bins=30) 
plot_grid(p12a,p12b,p12c,p12d,
          rows=2,
          cols=2,
          labels = c("a", "b", "c", "d"))

all_lip$c_lip_T<-(all_lip$c_lip) 
all_lip$dr_lip_T<-(all_lip$dr_lip) 
all_lip$hs_lip_T<-(all_lip$hs_lip) 

testing<-select(all_lip, RIL, c_lip_T, dr_lip_T, hs_lip_T)
write.table(testing, "../processed_data/lipid_transform.txt", sep="\t")
```

**CARBOHYDRATE:** DR is log; C is log; HS is untransformed

**LIPID:**  DR, C and HS untransformed;

**GLYCOGEN:** DR, C and HS all square root

**PROTEIN:** DR,C and HS is  untransformed

####FIX THIS 
#Complete transformations for each macromolec
```{r, include=FALSE}
carb_dr<-ggplot(data=all_carb, aes(x = (dr_carb)^(1/2))) +
  geom_histogram(bins=30) + labs(x="square root", y="carbohydrate", title="dietary restriction")
carb_c<-ggplot(data=all_carb, aes(x = log(c_carb))) +
  geom_histogram(bins=30)+labs(x="log", title="control",y=" ")
carb_hs<-ggplot(data=all_carb, aes(x = (hs_carb))) +
  geom_histogram(bins=30)  +labs(x="untransformed", title="high sugar",y=" ")

gly_dr<-ggplot(data=all_gly, aes(x = (dr_gly)^(1/2))) +
  geom_histogram(bins=30) + labs (x="square root", y="glycogen")
gly_c<-ggplot(data=all_gly, aes(x = (c_gly)^(1/2))) +
  geom_histogram(bins=30) +labs(x="square root",y=" ")
gly_hs<-ggplot(data=all_gly, aes(x = (hs_gly)^(1/2))) +
  geom_histogram(bins=30) +labs(x="square root",y=" ")

prot_dr<-ggplot(data=all_prot, aes(x = (dr_prot)^(1/2))) +
  geom_histogram(bins=30) + labs (x="square root", y="protein", title = "dietary restriction")
prot_c<-ggplot(data=all_prot, aes(x = log(c_prot))) +
  geom_histogram(bins=30) + labs (x="log",y=" ", title = "control")
prot_hs<-ggplot(data=all_prot, aes(x = (hs_prot))) +
  geom_histogram(bins=30) + labs (x="untransformed",y=" ", title= "high sugar")

lip_dr<-ggplot(data=all_lip, aes(x = (dr_lip))) +
  geom_histogram(bins=30) + labs (x="untransformed", y="lipid")
lip_c<-ggplot(data=all_lip, aes(x = (c_lip)^(1/2))) +
  geom_histogram(bins=30) + labs (x="square root",y=" ")
lip_hs<-ggplot(data=all_lip, aes(x = (hs_lip))) +
  geom_histogram(bins=30) + labs (x="untransformed",y=" ")


test_plot<-plot_grid(carb_dr,carb_c,carb_hs,gly_dr, gly_c,gly_hs,prot_dr,prot_c,prot_hs,lip_dr,lip_c,lip_hs,
          rows=4,
          cols=3)
#saved as pdf file under plots. 
```
#untransformed
```{r, include=FALSE}
carb_dr_ut<-ggplot(data=all_carb, aes(x = (dr_carb))) +
  geom_histogram(bins=30) + labs(x=" ", y="untransformed", title=" ")
carb_c_ut<-ggplot(data=all_carb, aes(x = (c_carb))) +
  geom_histogram(bins=30)+labs(x=" ", title=" ",y=" ")
carb_hs_ut<-ggplot(data=all_carb, aes(x = (hs_carb))) +
  geom_histogram(bins=30)  +labs(x=" ", title=" ",y=" ")

gly_dr_ut<-ggplot(data=all_gly, aes(x = (dr_gly))) +
  geom_histogram(bins=30) + labs (x=" ", y="untransformed")
gly_c_ut<-ggplot(data=all_gly, aes(x = (c_gly))) +
  geom_histogram(bins=30) +labs(x=" ",y=" ")
gly_hs_ut<-ggplot(data=all_gly, aes(x = (hs_gly))) +
  geom_histogram(bins=30) +labs(x=" ",y=" ")

prot_dr_ut<-ggplot(data=all_prot, aes(x = (dr_prot))) +
  geom_histogram(bins=30) + labs (x=" ", y="untransformed")
prot_c_ut<-ggplot(data=all_prot, aes(x = (c_prot))) +
  geom_histogram(bins=30) + labs (x=" ",y=" ")
prot_hs_ut<-ggplot(data=all_prot, aes(x = (hs_prot))) +
  geom_histogram(bins=30) + labs (x=" ",y=" ")

lip_dr_ut<-ggplot(data=all_lip, aes(x = (dr_lip))) +
  geom_histogram(bins=30) + labs (x=" ", y="untransformed")
lip_c_ut<-ggplot(data=all_lip, aes(x = (c_lip))) +
  geom_histogram(bins=30) + labs (x=" ",y=" ")
lip_hs_ut<-ggplot(data=all_lip, aes(x = (hs_lip))) +
  geom_histogram(bins=30) + labs (x=" ",y=" ")

carb_gly_TandUT_plots<-plot_grid(carb_dr,carb_c,carb_hs,
                     carb_dr_ut,carb_c_ut,carb_hs_ut,
                     gly_dr, gly_c,gly_hs,
                     gly_dr_ut, gly_c_ut,gly_hs_ut,
          rows=4,
          cols=3)
#SAVE THIS

prot_lip_TandUT_plots<-plot_grid(prot_dr,prot_c,prot_hs,
                     prot_dr_ut,prot_c_ut,prot_hs_ut,
                     lip_dr,lip_c,lip_hs,
                     lip_dr_ut,lip_c_ut,lip_hs_ut,
          rows=4,
          cols=3)

```
